\documentclass{article}

\usepackage{parskip}
\usepackage{fancyhdr}

\begin{document}
\lhead{04.05.10}
\textsc{choice} = \textsc{free will} --- \textsc{western religion} --- \textsc{christianity, judaism, islam.}

\textsc{no choice} = \textsc{fatalism} --- \textsc{eastern religion} --- \textsc{hinduism, buddhism, sikhism.}

Ethical goal for Christians is to become Christ-like.

How do Christians become Christ-like?  They follow the teachings of Jesus.


\textsl{Christianity and equality}

Are disabled people equal to able bodied people?

They should be and there are measures in place to make disabled people equal to able bodied people but they are not equal to able bodied people because they are not able bodied.

Disability: their world may be different but they should not be treated unequally.
Religion says yes they should be treated equally.

``\textsl{So there is no difference between Jews and Gentiles, between slaves and free people, between men and women; you are all in one union with Jesus Christ}''

\textbf{Christianity and Gender Equality}

St. Paul (Gal 3:28) may say that Vhristians should treat men and women equally but in practice they may differ because of:

\textsc{Law:}
\begin{itemize}
\item{in divorce cases children are nearly always placed with the mother}
\item{men do not have equal pension rights as widowers}
\end{itemize}

\textsc{Tradition:}
\begin{itemize}
\item{females are regarded as being more delicate than men}
\item{discrimination against women in sport}
\end{itemize}

\end{document}

