\documentclass{article}

\usepackage{parskip}
\usepackage{fancyhdr}

\begin{document}

\lhead{27.04.10}

How do you regard a still born child?
Should the baby have a religious funeral?

Religous perspectives on the nature and value of human life.

Religion or no religion can influence how you see human life, especially in terms of \textsc{abortion} and \textsc{euthanasia}.

You need to study Christianity and one religion of your choice in terms of ethics.

How you view \textsl{soul} can influence how you view \textsl{life}.

``God created human beings, making them to be like himself, male and female, and blessed them.'' Genesis 1 ,  27.

Humans have a unique nature and value in creation.  Human beings have sanctity (made holy) through a relationship with God.  < Not all people accept this, however.

Thomas Aquinas tried to show how animals and humans differed in status in terms of the relationship with God.  He linked \textsc{reasoning}, \textsc{ability} and \textsc{soul}.  Animals cannot \textsc{reason} therefore have no soul.  Special sanctity of life only remains between humans and God.  Because the sanctity of life is linked to human soul.  It refers being cremated in the likeness of God = a supernatural relationship which requires \textsc{faith}.

\subsubsection*{Christianity and the nature of human life}

\begin{tabular}{ | p{5cm} | p{5cm} | }
\hline
RC Church & Protestant Church\\ \hline
Original sin.  First sin.  Offense against God (sin) of disobedience.  St Augustin stated that the stain of original sin was passed through the sexual union of parents to children.  The Roman Catholic (RC) Church about 1000 years ago introduced infant baptism to wipe out the stain of original sin. & The Protestant view of the fall of Adam and Eve has meant that human beings are inherently weak or fallen.  They can overcome their weakness by turning to Christ and rejecting evil.  In general, human beings are generally imperfect and flawed and have to turn to Christ to be saved.  If you do sin (through e.g wrong ethical actions) you affect your human dignity and offend God.  This is why Christianity has a tending on \textsc{forgiveness}.\\ \hline
\end{tabular}


\end{document}
