\documentclass[11pt,a4paper]{article}

\usepackage{parskip,fancyhdr}
%\usepackage[T1]{fontenc}
%\usepackage{cmbright}

\setlength{\headheight}{15.2pt}
\pagestyle{fancy}

\lhead{Isabell Long}
\rhead{AS English Language \& Literature coursework}

\begin{document}

\title{Automonobiolography}
\author{Isabell Long}
\maketitle

Four hours down, two to go.  I am bored stiff---literally.  I am hungry, too.  I would get up, but to reach the restaurant coach on this train I have to go through first class and squeeze past the bar, just like on an East Coast train.  I have always maintained that East Coast, when you're on a packed train in the uncomfortable seats as opposed to the relatively comfortable grey ones, is the train equivalent of Ryanair, only without the cheap fares to match the quality of the experience.  This train is going from Penzance to Glasgow, taking just under nine hours, stopping at, um, around fourteen stations on the way.  I pity the people going all the way up.  Nearly ten hours on a train will be horrible, especially when on it from twenty past ten in the morning until quarter past eight at night!  Madness.  But I suppose people have their reasons, be they leisure- or work-related.  Just like I have my reasons for going to York.

York is amazing.  I come out of my shell.  It's as if I lead a double life!  Not that I know how long I can afford to keep going there, in holidays.  As a result, I am making the most of it while I can.  I've made a life for myself there.  It's quite weird.  I come out of my shell.  I have a fair few friends at the main University.  They all study Computer Science.  That's not a coincidence.  Real friends, not just acquaintances.  That distinction is important, don't you think?  I've never been one for having loads of friends, and have always got on better with adults than my peer group.  Weird.

`There's something changing inside of me, there's something changing inside of me\ldots'---sorry, that's my eccentricity coming through.  Those lyrics just came into my head.  Staring at one's reflection in a train window, when the sun's glare isn't too great, tends to do that to a person.  My feet are starting to go even more numb than they were, that dull ache.  And my coccyx is hurting now.  [\textsl{\textbf{pensive voice}}] Hmmm.  Ow.  [\textsl{\textbf{end pensive voice}}]  I knew I should have invested in one of those cushions!

[\textsl{\textbf{after three minutes}}] Ah, I'm back.  I just had to go for a walk as my legs were unhealthily stiff.  I bought chocolate, too.  Mmmm, chocolate.  So that I don't end up like a poor woman in a book I read a while ago, `The Trick is to Keep Breathing', who sits for hours so that her limbs go numb on purpose because it takes her mind off the depression, and doesn't eat, just survives on caffeine.  Ha.  No, that's not me.  [\textsl{\textbf{smiles}}]  Anyway.  I've always been a bit of a social misfit.  A self-confessed social misfit.  To tell the truth, sometimes I don't mind it.  I like my own company.  Yes, I am an only child.  Hence sitting here talking to myself, yet almost hoping that someone is recording and/or transcribing this for later.  Heh.

It all started in primary school.  We moved from the horror that was Southampton (I say that looking at it now---back then I was too young to remember), to Somerset.  My classmates took great delight in playing the [\textsl{\textbf{sarcastic voice}}] simply \textit{wonderful} [\textsl{\textbf{end sarcastic voice}}] game aptly named `pop the pig'.  The aim?  To squash the smallest girl in the class (me) with the heaviest (and most horrible) girl in the class.  You know when you laugh as an alternative to crying?  That.  So, as a result, none of the supervisors noticed that I was in immense discomfort.  Madness.  [\textsl{\textbf{shaky laugh}}]  Despite this, I still used to skip through the school gates every morning.  Bizarre.  I've never understood why.  It wasn't all bad---I managed to pluck up the courage to play a flute solo on stage in front of two hundred students, for example, after my Grade One exam, and joined the recorder group with my recorder and then my flute in Year 4---a year before anyone else was invited.  I remember I was so horribly nervous.  But then I used to feed positively off nerves, not like I do now and shake every time I have to do public speaking.  Life's changes and increases and decreases in confidence are interesting.  I don't play my flute any more though.  Sometimes, particularly around Christmas, I think how lovely it would be, but my priorities just aren't in the right place these days.

Oooh.  The train just pulled into the station at Leeds.  Not far to go.  I'm on the slow, cheap train, but with not far to go.  I still have time though.  Not that I want to have time.  I don't want to be in the cold---that'll set my fingers to purple mode again: the wind chill is like having air from a leafblower blown straight at your fingers, let alone the tingling when they eventually warm up, plus the fact that gloves hardly help---argh, Raynaud's---, but fresh air would be nice.

Fresh air?  What's wrong with me?  [\textsl{\textbf{laugh}}]  I'm a geek!  I don't like sunlight.  Well.  That's who I am now.  Again, my niche has been found.  An amazing community all around the country who congregate in London mostly, is my niche: Young Rewired State, and associated projects.  Ahhhh.  [\textsl{\textbf{positive sigh}}]  Apart from the presentations.  Although I have got better over the months!

Tada!  [\textsl{\textbf{train announce woman voice}}]  ``The next station is York.''  Finally!  I'm glad I got here safely---for some reason I had a morbid feeling that the train would break down!  I should go and find my stuff.  I hope no-one has mistaken my suitcase for theirs.  I blame my scatty brainedness for forgetting to put a label on it\ldots though that is only really required on planes these days.  Right, yes.  I'm off to enjoy my week of not sleeping \textsl{too} much---too much sleep makes people more tired!  I wonder if someone will be at the station to pick me up\ldots I can't remember what was decided, so I hope so!

\begin{center}
	\subsubsection*{Final word count (including stage directions): 1021 words.}
\end{center}

\end{document}
