\documentclass{article}

\usepackage{parskip}

\begin{document}

Animism --- use of animals as imagery for God\\
	--- worship of nature e.g a sacred mountain

One God, Brahma, the creator --- many many different forms (300 000)

Hindus believe that all life is sacret (holy to God) because all souls are immortal (cannot die).  They see God in all souls.  They also believe that the souls of all living creatures are one and the same as the eternal Brahma, therefore if you take life you hurt Brahma.

Plato believed in the transmigration of the soul and believed that the body is only a vehicule for the soul and is not important.

The sentient being is something that is capable of having senses.  Humans and animals are sentient.

The Hindu belief in the sacredness of life is connected with the idea that God is in everything and so to respect all life is to respect God.

Karma --- good things lead to pleasure in the next life, while best actions lead to suffering in the next life.

Fatalism --- accepting whatever happens to you
Stoic --- indifferent to pleasure or pain

The turning point for all human life is birth.  All life, especially the soul which is immortal, is sacred to Brahma.  The Hindu word for soul is \textsc{atman}.  To respect life is to respect God.  Ultimate Goal = union of the soul with Brahma to avoid reincarnation of the soul.

Free will --- having the right to make choices or decisions.

Some people say that in Britain we don't have total free will e.g. laws constrain us, we have to go to school until we are 16, we have to pay tax as adults...

Human beings have special value or worth because they are an instrument by which the soul can be liberated from the body and the cycle of rebirth --- reincarnation into human form makes this transition possible.


\textbf{Human nature and the value of human condition}\\
\textit{The purpose of human beings is to seek as an end pleasure through good action, individual liberation and liberation of society, welfare and righteousness}

Righteousness means having the correct relationship with God.  Hindus achieve this through doing the right thing to gain karma.

Hindus differ on the idea of whether the human soul is the same/equal to the soul of Brahma.  They all agree that the human form is one to escape from in order to have unity with God.

Liberation means freedom.

Welfare means caring for people.

The main ethic of Hunduism is always acting for the human family.

Hindu ethics are teleological (deals in the end result) which should be pleasure therefore evil is forbidden.

\end{document}
 

